An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-12 EconGrowth.Rnw
Now adding more file types to .gitignore to make collaboration easier.
A few small changes.
Deleted some bogus lines of text.
Now including tables of parameters for single-factor model fits.

<<single-factor functions, eval=TRUE>>=
####################
# Returns an nls single-factor model for the country and factor specified.
# factor should be one of "K", "L", "Q", "X", or "U".
##
singleFactorModel <- function(countryAbbrev, factor){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # We'll change the name of the desired column to "f"
  dataTable <- replaceColName(dataTable, factor, "f")
  # Now do the fit.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  mGuess <- 0.5
  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  modelSF <- nls(iGDP ~ exp(lambda*iYear) * f^m, 
                 data=dataTable, 
                 start = list(lambda=lambdaGuess, m=mGuess)
                 )
  return(modelSF)
}

@@

#################################################
# Calculates parameter estimates and confidence intervals
# for the single factor production function for a given a country.
#
# countryAbbrev is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
# factor is a string, one of "K", "L", "Q", "X", or "U"
#
# returns a data.frame of data for the Cobb-Douglas model. 
# First row is the +95% CI on all parameters
# Second row contains the parameter estimates
# Third row is the -95% CI on all parameters
# Each column has names: lambda and m corresponding to the parameters in the model.
##
singleFactorData <- function(countryAbbrev, factor){
  #First, check to see if we want useful work (U) AND one of the countries for which we don't have data.
  if (!haveDataSF(countryAbbrev, factor)){
    #Return a column of NAs if we don't have data for this factor
    nRows <- 3 # +95% CI, SF, and -95% CI.
    nCols <- 2 # lambda, m
    df <- as.data.frame(matrix(NA, ncol = nCols, nrow = nRows))
    colnames(df) <- c("lambda", "m")
    rownames(df) <- c("+95% CI", "SF", "-95% CI")
    return(df)
  }
  # We have a combination of country and factor for which we have data.
  modelSF <- singleFactorModel(countryAbbrev, factor)
  summarySF <- summary(modelSF) # Gives the nls summary table.
  ciSF <- confint(modelSF, level = ciLevel)  # Calculates confidence intervals for the SF model.
  dofSF <- summarySF$df[2] # Gives the degrees of freedom for the model.
  tvalSF <- qt(ciHalfLevel, df = dofSF)
#   print("")
#   print(paste("In singleFactorData:", countryAbbrev, factor))
#   print("summarySF:")
#   print(paste("    df:", summarySF$df))
#   print(paste("    sigma:", summarySF$sigma))
#   print("    coeffs: ")
#   print(summarySF$coefficients)
#   print("ciSF:")
#   print(ciSF)
  #Single factor model
  lambda <- as.numeric(coef(modelSF)["lambda"])
  m <- as.numeric(coef(modelSF)["m"])
  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estSF <- data.frame(lambda = lambda, m = m)
  row.names(estSF) <- "SF"
  # The [1] subscripts pick off the lower confidence interval
  lowerSF <- data.frame(lambda=ciSF["lambda","2.5%"], 
                        m=ciSF["m", "2.5%"])
  row.names(lowerSF) <- "-95% CI"
  # The [2] subscripts pick off the upper confidence interval
  upperSF <- data.frame(lambda=ciSF["lambda","97.5%"], 
                        m=ciSF["m", "97.5%"])
  row.names(upperSF) <- "+95% CI"
  # Now create the data for a table and return it
  dataSF <- rbind(upperSF, estSF, lowerSF)
  return(dataSF)
}

@@

Figure \ref{fig:CD_GDP_Lattice_Graph} compares predictions from the Cobb-Douglas models (without energy, with $Q$, and with $x$) to historical data.

@@

<<Print_Covariance_Tables, results='asis', eval=TRUE>>=
#I wish I could say this, but I get bogus "$US" strings appearing in the LaTeX output.
#lapply(countryAbbrevs, printCovarTable)
printCovarTable("US")
printCovarTable("UK")
printCovarTable("JP")
printCovarTable("CN")
printCovarTable("ZA")
printCovarTable("SA")
printCovarTable("IR")
printCovarTable("TZ")
printCovarTable("ZM")
@

@@

Tables \ref{} below provide tabular results for single-factor models.

<<SF_Params_Tables, results='asis', eval=TRUE>>=
# print(singleFactorParamsTable("Q"),
#      caption.placement="top", 
#      sanitize.colnames.function = identity, 
#      size="\\tiny", 
#      table.placement="H")
#printSingleFactorParamsTable("Q")
lapply(factors, printSingleFactorParamsTable)
@

@@

The Cobb-Douglas (with thermal energy, $q$) parameters are given in Table \ref{tab:CD_Parameters_With_Q}.

@@

Table \ref{tab:CD_Parameters_With_X} gives the parameters for the Cobb-Douglas model with exergy ($x$).

@@

The Cobb-Douglas (with useful work, $u$) parameters are given in Table \ref{tab:CD_Parameters_With_U}.

>>>>

<<single-factor functions, eval=TRUE>>=
####################
# Returns an nls single-factor model for the country and factor specified.
# factor should be one of "K", "L", "Q", "X", or "U".
##
singleFactorModel <- function(countryAbbrev, factor){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # We'll change the name of the desired column to "f"
  dataTable <- replaceColName(dataTable, factor, "f")
  # Now do the fit.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  # Some economies need different guesses
  if (countryAbbrev == "IR" && factor == "X"){
    mGuess <- 0.7
  } else {
    # This works for every other combination of country and factor.
    mGuess <- 0.5
  }
  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  modelSF <- nls(iGDP ~ exp(lambda*iYear) * f^m, 
                 data=dataTable, 
                 start = list(lambda=lambdaGuess, m=mGuess)
                 )
  return(modelSF)
}

@@

#################################################
# Calculates parameter estimates and confidence intervals
# for the single factor production function for a given a country.
#
# countryAbbrev is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
# factor is a string, one of "K", "L", "Q", "X", or "U"
#
# returns a data.frame of data for the Cobb-Douglas model. 
# First row is the +95% CI on all parameters
# Second row contains the parameter estimates
# Third row is the -95% CI on all parameters
# Each column has names: lambda and m corresponding to the parameters in the model.
##
singleFactorData <- function(countryAbbrev, factor){
  #First, check to see if we want useful work (U) AND one of the countries for which we don't have data.
  if (!haveDataSF(countryAbbrev, factor)){
    #Return a column of NAs if we don't have data for this factor
    nRows <- 3 # +95% CI, SF, and -95% CI.
    nCols <- 2 # lambda, m
    df <- as.data.frame(matrix(NA, ncol = nCols, nrow = nRows))
    colnames(df) <- c("lambda", "m")
    rownames(df) <- c("+95% CI", "SF", "-95% CI")
    return(df)
  }
  # We have a combination of country and factor for which we have data.
  modelSF <- singleFactorModel(countryAbbrev, factor)
  summarySF <- summary(modelSF) # Gives the nls summary table.
  ciSF <- confint(modelSF, level = ciLevel)  # Calculates confidence intervals for the SF model.
  dofSF <- summarySF$df[2] # Gives the degrees of freedom for the model.
  tvalSF <- qt(ciHalfLevel, df = dofSF)
  #Single factor model
  lambda <- as.numeric(coef(modelSF)["lambda"])
  m <- as.numeric(coef(modelSF)["m"])
  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estSF <- data.frame(lambda = lambda, m = m)
  row.names(estSF) <- "SF"
  # The [1] subscripts pick off the lower confidence interval
  lowerSF <- data.frame(lambda=ciSF["lambda","2.5%"], 
                        m=ciSF["m", "2.5%"])
  row.names(lowerSF) <- "-95% CI"
  # The [2] subscripts pick off the upper confidence interval
  upperSF <- data.frame(lambda=ciSF["lambda","97.5%"], 
                        m=ciSF["m", "97.5%"])
  row.names(upperSF) <- "+95% CI"
  # Now create the data for a table and return it
  dataSF <- rbind(upperSF, estSF, lowerSF)
  return(dataSF)
}

@@

Figure \ref{fig:CD_GDP_Lattice_Graph} compares predictions from the Cobb-Douglas models (without energy, with $q$, with $x$, and with $u$) to historical data.

@@

<<Print_Covariance_Tables, results='asis'>>=
#I wish I could say this, but I get bogus "$US" strings appearing in the LaTeX output.
#lapply(countryAbbrevs, printCovarTable)
printCovarTable("US")
printCovarTable("UK")
printCovarTable("JP")
printCovarTable("CN")
printCovarTable("ZA")
printCovarTable("SA")
printCovarTable("IR")
printCovarTable("TZ")
printCovarTable("ZM")
@

@@

The following tables provide tabular results for single-factor models.

<<SF_Params_Tables, results='asis'>>=
#I wish I could say this, but I get bogus K [1]” strings appearing in the LaTeX output.
#lapply(factors, printSingleFactorParamsTable)
printSingleFactorParamsTable("K")
printSingleFactorParamsTable("L")
printSingleFactorParamsTable("Q")
printSingleFactorParamsTable("X")
printSingleFactorParamsTable("U")
@

@@

The following tables provide tabular results for Cobb-Douglas models with various energy terms.

@@

@@
